---
title: "Fentanyl Dashboard"
output: 
  flexdashboard::flex_dashboard:
    theme: united
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(shiny)
library(janitor)
library(leaflet)
library(tidyverse)
library(sf)
library(tmap)
library(tmaptools)
library(geojsonio)
library(jsonlite)
library(plotly)
library(scales)
library(RColorBrewer)
library(httr)

# seizure_data <- read_csv("data/nationwide-drugs-fy21-fy24-jun.csv")
# scraped_data <- read_csv("final_df.csv")

#REMOVE THE TOKEN

# # The URLs to the raw CSV files
url1 <- "https://raw.githubusercontent.com/lukaslehmann-R/fentanyl/main/nationwide-drugs-fy21-fy24-jun.csv"
url2 <- "https://raw.githubusercontent.com/lukaslehmann-R/fentanyl/main/final_df.csv"



# Create GET requests with authentication
response1 <- GET(url1, authenticate(token, ""))
response2 <- GET(url2, authenticate(token, ""))


# Check if the requests were successful and read the content as CSV
if (response1$status_code == 200) {
  seizure_data <- read_csv(content(response1, "text"))
  print("nationwide-drugs-fy21-fy24-jun.csv loaded successfully")
  head(seizure_data)
} else {
  stop("Failed to retrieve nationwide-drugs-fy21-fy24-jun.csv: ", response1$status_code)
}

if (response2$status_code == 200) {
  scraped_data <- read_csv(content(response2, "text"))
  print("final_df.csv loaded successfully")
  head(scraped_data)
} else {
  stop("Failed to retrieve final_df.csv: ", response2$status_code)
}
# lab_data <- read_csv("data/dea_clan_lab_export.csv")

# unique(seizure_data$`Drug Type`)

```


```{r data prep}

seizure_data <- seizure_data %>%
  clean_names() %>%
  mutate(year = gsub(" \\(FYTD\\)", "", fy), 
  month = match(toupper(month_abbv), toupper(month.abb)),
  date = as.Date(paste0(year, "-", month, "-01")))

fent_seiz <- seizure_data %>%
  filter(drug_type == "Fentanyl")

years <- unique(year(fent_seiz$date))
months <- month.abb


states <- c(
  "Maryland", "Illinois", "Texas",
  "California", "Florida", "Louisiana",
  "New York", "Michigan", "Texas",
  "California", "Arizona", "Louisiana",
  "Washington", "Washington", "Vermont",
  "Texas", "Texas", "California",
  "Texas", "California", "Arizona",
  "Arizona", "California", "New York",
  "Texas", "Michigan", "Washington",
  "Montana", "Georgia", "Florida",
  "Massachusetts", "Texas", "New York",
  "Maine", "Puerto Rico", "Florida",
  "International", "Texas"
)

offices <- unique(fent_seiz$area_of_responsibility)

# Combine into a data frame
office_state_df <- data.frame(
  Office = offices,
  State = states,
  stringsAsFactors = FALSE
)

fent_w_states <- fent_seiz %>%
  left_join(office_state_df, by = c("area_of_responsibility" = "Office")) %>%
  clean_names() %>%
  filter(state != "International")

stategeojson <- geojson_read("https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json", what = "sp")

# stategeojson <- readRDS("stategeojson.rds")

# saveRDS(stategeojson, file = "stategeojson.rds")

# Summarize the data by state to get total sum_qty_lbs for each state
fent_by_state <- fent_w_states %>%
  group_by(state) %>%
  summarise(total_lbs = sum(sum_qty_lbs, na.rm = TRUE))

# Merge fentanyl data with geojson data
stategeojson@data <- left_join(stategeojson@data, fent_by_state, by = c("name" = "state"))


```

Fentanyl Seizures
============================================================================

Column {data-width=150}
-----------------------------------------------------------------------

### Sidebar

```{r}

selectInput("start_month", "Start Month", choices = months, selected = "Jan")
selectInput("start_year", "Start Year", choices = years, selected = min(years))
selectInput("end_month", "End Month", choices = months, selected = "Dec")
selectInput("end_year", "End Year", choices = years, selected = max(years))

actionButton("update", "Update Map")

```

Column {data-width=690}
-----------------------------------------------------------------------

### Chart A

```{r}

leafletOutput("seizure_map", width = "100%", height = "600px")

observeEvent(input$update, {
  
  start_date <- as.Date(paste(input$start_year, match(input$start_month, months), "1", sep = "-"))
  end_date <- as.Date(paste(input$end_year, match(input$end_month, months), "1", sep = "-"))
  
  # Filter the data based on the selected date range
  filtered_data <- fent_w_states %>%
    filter(date >= start_date & date <= end_date)
  
  # Summarize the data by state based on the filtered data
  fent_by_state_filtered <- filtered_data %>%
    group_by(state) %>%
    summarise(total_lbs = sum(sum_qty_lbs, na.rm = TRUE))
  
  # Clean state names in GeoJSON data and filtered data for consistency
  stategeojson@data$name <- tolower(trimws(stategeojson@data$name))
  fent_by_state_filtered$state <- tolower(trimws(fent_by_state_filtered$state))
  
  # Remove any existing total_lbs column in stategeojson@data to avoid conflicts during the join
  stategeojson@data$total_lbs <- NULL
  
  # Merge filtered fentanyl data with geojson data
  stategeojson@data <- left_join(stategeojson@data, fent_by_state_filtered, by = c("name" = "state"))

  # Replace NA values in total_lbs with 0 for states with no data
  stategeojson@data$total_lbs[is.na(stategeojson@data$total_lbs)] <- 0

  # Apply a logarithmic transformation to total_lbs for better visualization
  stategeojson@data$log_total_lbs <- log10(stategeojson@data$total_lbs + 1)

  # Define bins based on the actual seizure values
  bins <- c(0, 1, 10, 100, 1000, 5000, max(stategeojson@data$total_lbs, na.rm = TRUE))
  
  # Set up a color palette based on the log-transformed values
  pal <- colorBin(palette = "YlOrRd", domain = stategeojson@data$total_lbs, bins = bins, na.color = "transparent")
  
  leafletProxy("seizure_map", data = stategeojson) %>%
  clearShapes() %>%
  clearControls() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    fillColor = ~pal(total_lbs),  # Color based on the log-transformed values
    color = "#BDBDC3",
    weight = 1,
    fillOpacity = 0.7,
    # Use format() to add commas to the rounded total_lbs values
    label = ~paste0(name, ": ", format(round(total_lbs), big.mark = ","), " lbs"),
    layerId = ~name,
    highlightOptions = highlightOptions(weight = 2, color = "#666", fillOpacity = 0.7, bringToFront = TRUE)) %>%
  addLegend(pal = pal, values = ~total_lbs, bins = bins, opacity = 0.7,
            title = "Total Fentanyl Seizures (lbs)", position = "bottomright")


})

output$seizure_map <- renderLeaflet({
  leaflet(stategeojson) %>%
    addProviderTiles("CartoDB.Positron") %>%
    setView(lng = -96, lat = 37.8, zoom = 3.5) # Adjusted zoom level
})

```

Column {data-width=600}
-----------------------------------------------------------------------

### Chart B

```{r}

selected_state <- reactiveVal(NULL)

observeEvent(input$seizure_map_shape_click, {
  # Extract the clicked state's ID
  clicked_state <- input$seizure_map_shape_click$id
  # Set the clicked state to the reactive value
  selected_state(clicked_state)
})

output$a <- renderPlotly({
  req(selected_state())
  req(input$update)
  
  state_name <- selected_state()
  
  start_date <- as.Date(paste(input$start_year, match(input$start_month, month.abb), "1", sep = "-"))
  end_date <- as.Date(paste(input$end_year, match(input$end_month, month.abb), "1", sep = "-"))
  
  fent_w_states <- fent_w_states %>%
    mutate(state = tolower(state))
  
  # Filter the data based on the selected date range and state
  filtered_data1 <- fent_w_states %>%
    filter(date >= start_date & date <= end_date & state == state_name) %>%
    group_by(area_of_responsibility, date) %>%
    summarise(total_seizure = sum(sum_qty_lbs, na.rm = TRUE)) %>%
    ungroup()
  
  
  # Check if filtered data is empty
  if(nrow(filtered_data1) == 0) {
    showNotification("No data available for the selected state and date range.", type = "warning")
    return(NULL)
  }
  
  gg1 <- ggplot(filtered_data1, aes(x = date, y = round(total_seizure, 0), color = area_of_responsibility, 
                                    group = area_of_responsibility, text = paste("Date: ", date, "<br>Total Fentanyl Seizure (lbs): ", total_seizure))) +
    geom_line() +
    geom_point() +
    labs(title = paste("Fentanyl Seizure by Office in", str_to_title(state_name)),
         x = "Date",
         y = "Total Seizure",
         color = "Office") +
    theme_minimal()
  
  ggplotly(gg1, tooltip = "text")
})

# Output the plotly chart
plotlyOutput("a", width = "80%", height = "310px")

```

### Chart C

```{r}

selected_state <- reactiveVal(NULL)

# Observe the map click event and update the selected state
observeEvent(input$seizure_map_shape_click, {
  clicked_state <- input$seizure_map_shape_click$id  # Make sure the map input ID is correct
  selected_state(clicked_state)
})

# Render the Plotly chart
output$b <- renderPlotly({
  # Ensure that a state is selected and the update button is clicked
  req(selected_state())
  req(input$update)

  state_name <- selected_state()  # Use the selected state

  # Convert start and end dates from user input
  start_date <- as.Date(paste(input$start_year, match(input$start_month, month.abb), "1", sep = "-"))
  end_date <- as.Date(paste(input$end_year, match(input$end_month, month.abb), "1", sep = "-"))
  
  state_name <- tolower(state_name)  # Convert to lowercase to match data
  
  # Convert state names to lowercase in the data
  fent_w_states <- fent_w_states %>%
    mutate(state = tolower(state))
  
  # Filter the data based on the selected date range and state
  filtered_data2 <- fent_w_states %>%
    filter(date >= start_date & date <= end_date & state == state_name) %>%
    group_by(area_of_responsibility) %>%
    summarise(total_seizure = sum(sum_qty_lbs, na.rm = TRUE)) %>%
    ungroup()
  
  # Create tooltip text
  filtered_data2$tooltip_text <- paste("Authority: ", filtered_data2$area_of_responsibility, 
                                     "<br>Total Fentanyl Seizure (lbs): ", 
                                     round(filtered_data2$total_seizure, 0))
  
  # Create the ggplot object
  gg2 <- ggplot(filtered_data2, aes(x = reorder(area_of_responsibility, total_seizure), 
                                  y = round(total_seizure, 0), 
                                  fill = area_of_responsibility, 
                                  group = area_of_responsibility, 
                                  text = tooltip_text)) +
    geom_col() +
    coord_flip() +
    labs(title = paste("Fentanyl Seizure by Office in", str_to_title(state_name)),
         x = "Authority",
         y = "Total Seizure",
         fill = "Office") +  # Changed 'color' to 'fill'
    theme_minimal()
  
  # Render the chart with tooltips
  ggplotly(gg2, tooltip = "text")
})

# Output the Plotly chart in the UI
plotlyOutput("b", width = "80%", height = "310px")


```


CBP Contact Info
============================================================================

Column {data-height=100}
-----------------------------------------------------------------------

### Select a City

```{r}

scraped_data <- scraped_data %>%
  arrange(city_name)

# # Clean address_info in the dataframe
# scraped_data <- scraped_data %>%
#   mutate(address_info = gsub("([A-Za-z0-9]+)([A-Z][a-z]+,)", "\\1\n\\2", address_info))  # Adding a newline before city and state
# 
# scraped_data <- scraped_data %>%
#   mutate(address_info = gsub("- -", "", address_info))
# 
# # Clean address_info in the dataframe
# scraped_data <- scraped_data %>%
#   mutate(address_info = gsub("Suite", "\nSuite", address_info))  # Add a newline before "Suite"
# 
# # Another option is to add a space before United States
# scraped_data <- scraped_data %>%
#   mutate(address_info = gsub("United States", "\nUnited States", address_info))  # Add a newline before "United States"

# City selection input
selectInput("selected_city", "", choices = unique(scraped_data$city_name), selected = NULL)

# Action button to trigger the display of contact info
actionButton("go_button", "Go")

verbatimTextOutput("contact_info")

# # Server-side logic
observeEvent(input$go_button, {
  # Triggered when the user clicks the "Go" button

  # Reactive expression to filter the contact info based on selected city
  selected_contact_info <- reactive({
    req(input$selected_city)  # Ensure a city is selected
    scraped_data %>%
      filter(city_name == input$selected_city) %>%
      select(clean2, phone_numbers)
  })

  # Render the contact info when the button is clicked
  output$contact_info <- renderText({
    info <- selected_contact_info()

    # Format the output with address and phone number
    paste0("Address: ", info$clean2, "\nPhone Number: ", info$phone_numbers)
  })

})



```


About - READ ME
============================================================================

Hello C4ADS! I made this dashboard to quickly demonstrate a few of my R programming skills. Please explore this dashboard constructed from publicly available sources. I've chosen this topic because counternarcotics research is featured on your website.

<br>

This is a relatively simple dashboard, but increasing levels of complexity could be added depending on what is needed by the user. The general idea behind this dashboard is that (if the data is regularly updated) it could be used by someone interested in tracking trends in fentanyl seizures as a launching point for more in-depth investigation. 

<br>

The first tab provides a map of fentanyl seizures around the country. The scale on the map is a log scale so that the differences between state seizure numbers can be seen more easily. The second tab provides the contact information for the authorities seizing the fentanyl to make asking CBP questions related to these seizures easier. There were no email addresses available. 

<br>

The fentanyl seizure data comes from U.S. Customs and Border Protection. I scraped the contact info from their site using R.


```{r}
```

